---
title: "Financial Trading Exercise"
author: "dearbharat"
date: "10/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Quant Strat

We will how to construct a basic trading strategy in quantstrat, R’s industrial-strength backtesting platform developed by Brian Peterson, Director of Algorithmic Trading at DV Trading. You’ll build a trading strategy in quantstrat from start to finish, including the code to set up a new strategy, as well as the design of indicators, signals, and rules for your strategy. By the end, you’ll be ready to design and implement your own trading strategies directly in R!

quantstrat is currently only available on GitHub. If you want to install it on your own machine, you first need to remotes package.

```{r installation, include=FALSE}
# install.packages("devtools") # if not installed
# install.packages("FinancialInstrument") #if not installed
# install.packages("PerformanceAnalytics") #if not installed

# next install blotter from GitHub
devtools::install_github("braverock/blotter", force = TRUE)
# next install quantstrat from GitHub
devtools::install_github("braverock/quantstrat", force = TRUE)
```

### Mechanics of trading

**Trend trading** (also divergence or momentum), which is a bet that a quantity, such as a price, will keep moving in its current direction.

**Reversion trading** (also convergence, cycle, or oscillation), which is a bet that a quantity, such as a price, will reverse.

### What is overfitting

When developing a trading system, a major pitfall that can creep into system development is the desire to find a strategy that worked phenomenally in the past. This is known as overfitting. Research by leading authors in the quantitative field has shown that not only is an overfitted system unlikely to generate profits in the future, but also that its performance can lead to losses.


### Getting financial data

```{r getdata, message=FALSE, warning=FALSE}
# Get SPY from yahoo
library(tidyquant)
getSymbols("SPY", 
           from = "2000-01-01", 
           to = "2016-06-30", 
           src = "yahoo", 
           adjust = TRUE)

```

```{r}
# Plot the closing price of SPY
plot(Cl(SPY))

```


### Adding indicators to financial data

Adding a moving average to financial data

```{r}
# Add a 200-day SMA using lines()
lines(SMA(Cl(SPY), n = 200), col = "red")

```

## quantstrat strategies

```{r, message=FALSE}
# Load the quantstrat and quantmod
library(quantstrat)
library(quantmod)

```

### Initial settings
```{r}
# Create initdate, from, and to strings
initdate <- "1999-01-01"
from <- "2003-01-01"
to <- "2015-12-31"

# Set the timezone to UTC
Sys.setenv(TZ = "UTC")

# Set the currency to USD 
currency("USD")

# Define your trade size and initial equity
tradesize <- 100000
initeq <- 100000

# Define the names of your strategy, portfolio and account
strategy.st <- "firststrat"
portfolio.st <- "firststrat"
account.st <- "firststrat"

# Remove the existing strategy if it exists
# rm.strat(strategy.st)

# Initialize the portfolio
initPortf(portfolio.st, symbols = "SPY", initDate = initdate, currency = "USD")

# Initialize the account
initAcct(account.st, portfolios = portfolio.st, initDate = initdate, currency = "USD", initEq = initeq)

# Initialize the orders
initOrders(portfolio.st, initDate = initdate)

# Store the strategy
strategy(strategy.st, store = TRUE)

```

### Retrieve Stock Price Data
```{r , warning=FALSE}
# Retrieve SPY from yahoo
getSymbols("SPY", from = from, to = to, src = "yahoo", adjust = TRUE)
# Use stock() to initialize SPY and set currency to USD
stock("SPY", currency = "USD")

```

## Indicators
Introduction to indicators

The SMA and RSI functions

```{r}
# Create a 200-day SMA
spy_sma <- SMA(x = Cl(SPY), n = 200)

# Create an RSI with a 3-day lookback period
spy_rsi <- RSI(price = Cl(SPY), n = 3)

```


Visualize a trend indicator

```{r}
# Plot the closing prices of SPY
plot(Cl(SPY))

# Overlay a 200-day SMA
lines(SMA(Cl(SPY), n = 200), col = "red")


```

Visualize an reversion indicator

```{r}
# Plot the closing price of SPY
plot(Cl(SPY))

# Plot the RSI 2
plot(RSI(Cl(SPY), n = 2))

```


### Indicator mechanics

```{r}
# Add a 200-day SMA indicator to strategy.st
add.indicator(strategy = strategy.st, 
              
# Add the SMA function
name = "SMA", 
              
# Create a lookback period
arguments = list(x = quote(Cl(mktdata)), n = 200), 
              
# Label your indicator SMA200
label = "SMA200")

```


### Implementing an indicator - II

```{r}
# Add a 50-day SMA indicator to strategy.st
add.indicator(strategy = strategy.st, 
              
              # Add the SMA function
              name = "SMA", 
              
              # Create a lookback period
              arguments = list(x = quote(Cl(mktdata)), n = 50), 
              
              # Label your indicator SMA50
              label = "SMA50")


```
### Implementing an indicator - III
```{r}
# Add an RSI 3 indicator to strategy.st
add.indicator(strategy = strategy.st, 
              
              # Add the RSI 3 function
              name = "RSI", 
              
              # Create a lookback period
              arguments = list(price = quote(Cl(mktdata)), n = 3), 
              
              # Label your indicator RSI_3
              label = "RSI_3")

```


### Code your own indicator

```{r}
# Write the calc_RSI_avg function
calc_RSI_avg <- function(price, n1, n2) {
  
  # RSI 1 takes an input of the price and n1
  RSI_1 <- RSI(price = price, n = n1)
  
  # RSI 2 takes an input of the price and n2
  RSI_2 <- RSI(price = price, n = n2)
  
  # RSI_avg is the average of RSI_1 and RSI_2
  x <- (RSI_1 + RSI_2)/2
  
  # Your output of RSI_avg needs a column name of"RSI_avg"
  colnames(x) <- "RSI_avg"
  return(x)
}

# Add this function as RSI_3_4 to your strategy with n1 = 3 and n2 = 4
add.indicator(strategy.st, name = "calc_RSI_avg", arguments = list(price=quote(Cl(mktdata)), n1 = 3, n2 = 4), label = "RSI_3_4")

```


### Code your own indicator - II

```{r}
# Declare the DVO function
DVO <- function(HLC, navg = 2, percentlookback = 126) {
  
  # Compute the ratio between closing prices to the average of high and low
  ratio <- Cl(HLC)/((Hi(HLC) + Lo(HLC))/2)
  
  # Smooth out the ratio outputs using a moving average
  avgratio <- SMA(ratio, n = navg)
  
  # Convert ratio into a 0-100 value using runPercentRank()
  out <- runPercentRank(avgratio, n = percentlookback, exact.multiplier = 1) * 100
  colnames(out) <- "DVO"
  return(out)
}

# Add the DVO indicator to your strategy
add.indicator(strategy = strategy.st, name = "DVO",
              arguments = list(HLC = quote(HLC(mktdata)), navg = 2, percentlookback = 126),
              label = "DVO_2_126")

```
```{r}
# Use applyIndicators to test out your indicators
test <- applyIndicators(strategy = strategy.st, mktdata = OHLC(SPY))


```

```{r}
# Subset your data between Sep. 1 and Sep. 5 of 2013
test_subset <- test["2013-09-01/2013-09-05"]

```


## Signals
A signal is an interaction of market data with indicators, or indicators with other indicators, which tells you whether you may wish to buy or sell an asset. Signals can be triggered for a variety of reasons. For example, a signal may be triggered by a shorter lookback moving average going from less than to greater than a longer lookback moving average. Another signal may be triggered when an oscillator goes from being above a certain set quantity to below, and so on.

You will see various ways in which indicators interact with each other. To keep thing simple, you will remove all of the RSI indicators and stick to the DVO (David Varadi’s Oscillator) indicator you implemented earlier.

Exercise: Subset test between September 10th, 2010, and October 10th, 2010, using test["YYYY-MM-DD/YYYY-MM-DD"]

Is SMA50 greater than or less than SMA200 on September 20?


You will manually do a sigThreshold-type evaluation without yet calling the signal. sigThreshold is a signal threshold argument which assesses whether or not a value is above or below a certain static quantity. you will apply simple moving averages and the DVO you implemented earlier.

Exercise: Subset test between September 10th, 2010, and October 10th, 2010, using test["YYYY-MM-DD/YYYY-MM-DD"]

Is DVO greater or smaller than 20 on September 30?

### sigComparison and sigCrossover

```{r}
# Add a sigComparison which specifies that SMA50 must be greater than SMA200, call it longfilter
add.signal(strategy.st, name = "sigComparison", 
           
           # We are interested in the relationship between the SMA50 and the SMA200
           arguments = list(columns = c("SMA50", "SMA200"), 
                            
                            # Particularly, we are interested when the SMA50 is greater than the SMA200
                            relationship = "gt"),
           
           # Label this signal longfilter
           label = "longfilter")

```


### Using sigCrossover

```{r}
# Add a sigCrossover which specifies that the SMA50 is less than the SMA200 and label it filterexit
add.signal(strategy.st, name = "sigCrossover",
           
           # We're interested in the relationship between the SMA50 and the SMA200
           arguments = list(columns = c("SMA50", "SMA200"),
                            
                            # The relationship is that the SMA50 crosses under the SMA200
                            relationship = "lt"),
           
           # Label it filterexit
           label = "filterexit")
```

### Using sigThreshold - I

```{r}
# Implement a sigThreshold which specifies that DVO_2_126 must be less than 20, label it longthreshold
add.signal(strategy.st, name = "sigThreshold", 
           
           # Use the DVO_2_126 column
           arguments = list(column = "DVO_2_126", 
                            
                            # The threshold is 20
                            threshold = 20, 
                            
                            # We want the oscillator to be under this value
                            relationship = "lt", 
                            
                            # We're interested in every instance that the oscillator is less than 20
                            cross = FALSE), 
           
           # Label it longthreshold
           label = "longthreshold")
```

### Using sigThreshold() - II

```{r}
# Add a sigThreshold signal to your strategy that specifies that DVO_2_126 must cross above 80 and label it thresholdexit
add.signal(strategy.st, name = "sigThreshold", 
           
           # Reference the column of DVO_2_126
           arguments = list(column = "DVO_2_126", 
                            
                            # Set a threshold of 80
                            threshold = 80, 
                            
                            # The oscillator must be greater than 80
                            relationship = "gt", 
                            
                            # We are interested only in the cross
                            cross = TRUE), 
           
           # Label it thresholdexit
           label = "thresholdexit")
```

### Using sigFormula()
```{r}
# Create your dataset: test
test_init <- applyIndicators(strategy.st, mktdata = OHLC(SPY))
test <- applySignals(strategy = strategy.st, mktdata = test_init)
```

**Combining signals - I**

You’ll want to create a signal when BOTH longfilter and longthreshold are equal to 1. 

Exercise: Have a look at test on October 8, 2013. Are longfilter and longthreshold both equal to 1 on that date?

**Combining signals - II**


```{r}
# Add a sigFormula signal to your code specifying that both longfilter and longthreshold must be TRUE, label it longentry
add.signal(strategy.st, name = "sigFormula",         
           # Specify that longfilter and longthreshold must be TRUE
           arguments = list(formula = "longfilter & longthreshold", 
                            
                            # Specify that cross must be TRUE
                            cross = TRUE),
           
           # Label it longentry
           label = "longentry")

```
     
## Rules
Using add.rule() to implement an exit rule
```{r}
# Fill in the rule's type as exit
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Specifying sigcol in add.rule()**

```{r}
# Fill in the sigcol argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Specifying sigval in add.rule()**
```{r}
# Fill in the sigval argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Specifying orderqty in add.rule()**
```{r}
# Fill in the orderqty argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")
```

**Specifying ordertype in add.rule()**
```{r}
# Fill in the ordertype argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Specifying orderside in add.rule()**
```{r}
# Fill in the orderside argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Specifying replace in add.rule()**
```{r}
# Fill in the replace argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "thresholdexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Specifying prefer in add.rule()**

```{r}
# Fill in the prefer argument in add.rule()
add.rule(strategy.st, name = "ruleSignal", 
         arguments = list(sigcol = "thresholdexit", sigval = TRUE, orderqty = "all", 
                        ordertype = "market", orderside = "long", 
                        replace = FALSE, prefer = "Open"), 
         type = "exit")

```

**Using add.rule() to implement an entry rule**
```{r}
# Create an entry rule of 1 share when all conditions line up to enter into a position
add.rule(strategy.st, name = "ruleSignal", 
         
         # Use the longentry column as the sigcol
         arguments = list(sigcol = "longentry", 
                        
                        # Set sigval to TRUE
                        sigval = TRUE, 
                        
                        # Set orderqty to 1
                        orderqty = 1,
                        
                        # Use a market type of order
                        ordertype = "market", 
                        
                        # Take the long orderside
                        orderside = "long", 
                        
                        # Do not replace other signals
                        replace = FALSE, 
                        
                        # Buy at the next day's opening price
                        prefer = "Open"), 
         
         # This is an enter type rule, not an exit
         type = "enter")

```

### Order sizing functions

**Implementing a rule with an order sizing function**
```{r}
# Add a rule that uses an osFUN to size an entry position
add.rule(strategy = strategy.st, name = "ruleSignal",
         arguments = list(sigcol = "longentry", sigval = TRUE, ordertype = "market",
                          orderside = "long", replace = FALSE, prefer = "Open",
                          
                          # Use the osFUN called osMaxDollar
#                          osFUN = osMaxDollar,
                          
                          # The tradeSize argument should be equal to tradesize (defined earlier)
                          tradeSize = tradesize,
                          
                          # The maxSize argument should be equal to tradesize as well
                          maxSize = tradesize),
         type = "enter")

```

## Analyzing your strategy

Running your strategy
```{r}
# Use applyStrategy() to apply your strategy. Save this to out
out <- applyStrategy(strategy = strategy.st, portfolios = portfolio.st)

```
```{r}
# Update your portfolio (portfolio.st)
updatePortf(portfolio.st)

```

```{r}
daterange <- time(getPortfolio(portfolio.st)$summary)[-1]

# Update your account (account.st)
updateAcct(account.st, daterange)

```

```{r}
updateEndEq(account.st)

```

**Profit factor**
```{r}
# Get the tradeStats for your portfolio
tstats <- tradeStats(Portfolios = portfolio.st)
# Print the profit factor
tstats$Profit.Factor
```

### Percent positive

While profit factor is one important statistic, it may be heavily influenced by only a few good trades. The percent positive statistic lets you know how many of your trades were winners. A trading system based on oscillation trading will likely have a high percentage of winners. This is certainly a statistic you should look for in your own trade statistics.

Exercise: The trading statistics object you created in the last exercise (tstats) has been preloaded into your workspace. Examine it. What is the percent positive statistic?

### Visualizing your strategy
Using chart.Posn()
```{r}
# Use chart.Posn to view your system's performance on SPY
chart.Posn(Portfolio = portfolio.st, Symbol = "SPY")


```

### Adding an indicator to a chart.Posn() chart
```{r}

# Compute the SMA50
sma50 <- SMA(x = Cl(SPY), n = 50)

# Compute the SMA200
sma200 <- SMA(x = Cl(SPY), n = 200)

# Compute the DVO_2_126 with an navg of 2 and a percentlookback of 126
dvo <- DVO(HLC = HLC(SPY), navg = 2, percentlookback = 126)

# Recreate the chart.Posn of the strategy from the previous exercise
chart.Posn(Portfolio = portfolio.st, Symbol = "SPY")

# Overlay the SMA50 on your plot as a blue line
add_TA(sma50, on = 1, col = "blue")

# Overlay the SMA200 on your plot as a red line
add_TA(sma200, on = 1, col = "red")

# Add the DVO_2_126 to the plot in a new window
add_TA(dvo)
```


### Additional analytics

Cash Sharpe ratio

When working with cash profit and loss statistics, quantstrat offers a way to compute a Sharpe ratio not just from returns, but from the actual profit and loss statistics themselves. A Sharpe ratio is a metric that compares the average reward to the average risk taken. Generally, a Sharpe ratio above 1 is a marker of a strong strategy.

In this exercise, you will see that because of trading P&L (profit and loss), one can compute a Sharpe ratio based on these metrics. The code below can be used to compute the Sharpe ratio based off of P&L. Copy the code in the console. In what range is the Sharpe ratio you obtain?

```{r}
# portpl <- .blotter$portfolio.firststrat$summary$Net.Trading.PL SharpeRatio.annualized(portpl, geometric=FALSE)

```

### Returns Sharpe ratio in quantstrat
```{r}
# Get instrument returns
instrets <- PortfReturns(portfolio.st)

# Compute Sharpe ratio from returns
SharpeRatio.annualized(instrets, geometric = FALSE)

```