---
title: "HTML Widget"
author: "dearbharat"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GDP Choropleth

The `leaflet` library allows us to create interactive maps, including shaded-area maps (choropleth) like the one below.

```{r gdp-choropleth, fig.cap="Choropleth showing per country GDP (Total \\$ in 2009)"}
library("tidyverse")
library("gapminder")
library("glue")
library("rnaturalearth")
library("sf")
library("leaflet")
library("leaflet.extras")

countries_sf <- countries110 %>%
  st_as_sf()

pop_palette <- colorNumeric("viridis",
                            domain = countries_sf$gdp_md_est)

countries_sf %>%
  filter(! name == "Antarctica") %>%
  leaflet() %>%
  # addProviderTiles(providers$Esri.WorldShadedRelief) %>%
  addPolygons(fillColor = ~pop_palette(gdp_md_est),
              color = "black",
              weight = 1,
              opacity = 1,
              fillOpacity = 0.7,
              popup = ~glue("Country: {name}", "<br>",
                           "Population: {pop_est}",
                           "<br>",
                           "GDP: {gdp_md_est}")) %>%
  addLegend(pal = pop_palette,
            values = ~gdp_md_est,
            opacity = 0.7,
            title = "GDP in $ (2009)", position = "bottomleft") %>%
  setMapWidgetStyle(style = list(background = "#aacbff"))


```


```{r setup1, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	fig.pos = "h",
	out.extra = ""
)
```

# Population Growth by Continent

This example demonstrates how to embed `htmlwidgets` in HTML formats. Note that we've demonstrated how `htmlwidgets` can be used with captions.

We need the following packages for this document:

```{r}
library("gapminder")
library("tidyverse")
library("highcharter")
library("lubridate")
library("glue")
library("leaflet")
library("leaflet.extras")
library("sf")
library("rnaturalearth")
```

To prevent issues with \LaTeX being clever with image placements, the command `\newpage` has been used below.

\newpage


The visualisation below was created with `highcharter`. The `highcharter` library provides the ability to create a wide range of beautiful, professional looking charts. 

```{r gapminder-pop-growth, fig.cap="Continental population growth (Source: Gapminder)"}
gapminder_continents <- gapminder %>%
  group_by(year, continent) %>%
  summarise_if(is.numeric, funs(round(mean(as.numeric(.))))) %>%
  ungroup() %>%
  mutate(year = ymd(glue("{year}-12-31")))
  
highchart(type = "stock") %>%
  hc_add_series(data = gapminder_continents,
                type = "line",
                hcaes(x = year,
                      y = pop,
                      group = continent)) %>%
  hc_plotOptions(series = list(marker = list(enabled = TRUE))) %>%
  hc_tooltip(xDateFormat = "%Y") %>%
  hc_legend(enabled = TRUE, reverse = TRUE)
```


# Leaflet

[Leaflet](http://rstudio.github.io/leaflet/)

Leaflet is a JavaScript library for creating dynamic maps that support panning and zooming along with various annotations like markers, polygons, and popups.

```{r}
library(leaflet)
pal <- colorQuantile("YlOrRd", NULL, n = 8)
orstationc <- read.csv("orstationc.csv")
leaflet(orstationc) %>% 
  addTiles() %>%
  addCircleMarkers(color = ~pal(tann))
```

# Dygraphs
```{r}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

