---
title: "Portfolio Analysis"
author: "dearbharat"
date: "10/2/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Portfolio Analysis

Diversify investment -> porfolio -> increase return’s expectation and reduce the risk. Hence, we need to test our investment strategy, it is called: backtesting. Backtesting are tested using historical data. We also need to use an online performance monitoring.

## Import Data
```{r}
library(tidyquant)
ko <- getSymbols("ko", from = "2003-01-01", to = "2016-08-30", auto.assign = FALSE)
pep <- getSymbols("pep", from = "2003-01-01", to = "2016-08-30", auto.assign = FALSE)
head(cbind(ko, pep))
```

## Data Manipulation

First, we define ko_pep as the ratio expressing the value of the share price of the Coca Cola company in terms of the share price of PepsiCo.

```{r}
ko <- Ad(ko)
pep <- Ad(pep)
head(cbind(ko, pep))
```

```{r}
ko_pep <- ko/pep
head(ko_pep)
```

## Data Exploration
```{r}
# Make a time series plot of ko_pep
plot.zoo(ko_pep)
  
# Add as a reference, a horizontal line at 1
abline(h=0.5)
```

## Portfolio Strategy Question
How to decide between these two companies? Coca-cola or Pepsi? 30% KO and 70% PEP, or just 100% PEP just because you like Pepsi more?

Several strategies available to choose: 
1. Put all-in on one basket <- Noobs only! 
2. Equal weighting 
3. Market cap weighting 
4. Mean and variance optimization <- Our goal!

Return.calculate(): calculate the individual asset return Return.portfolio(): calculate the porfolio return

```{r}
library(PerformanceAnalytics)
ko_ret <- Return.calculate(ko)
pep_ret <- Return.calculate(pep)
head(cbind(ko_ret, pep_ret))
```

```{r}
# Remove the first row (NA)
ko_ret <- ko_ret[(-1),]
pep_ret <- pep_ret[(-1),]
```

For portfolio weights, there are options: 
1. Set initial weight at initial date and never change it. So, the weight of individual stock price in the portfolio will constantly changing because the changes of the price itself. 
2. Dynamic approach -> we constantly adjusting the weight of the individual asset.

Let’s try to build both models!

```{r}
# Combine both returns into one variable
rets <- cbind(ko_ret, pep_ret)

# Create the weights, we divide our money half to each of the asset
eq_weights <- c(0.5, 0.5)

# Create a portfolio using buy and hold
pf_bh <- Return.portfolio(R = rets, weights = eq_weights, verbose = TRUE)

# Create a portfolio rebalancing monthly 
pf_rebal <- Return.portfolio(R = rets, weights = eq_weights, rebalance_on = "months", verbose = TRUE)

# Plot the time-series
par(mfrow = c(2, 1), mar = c(2, 4, 2, 2))
plot.zoo(pf_bh$returns)
plot.zoo(pf_rebal$returns)
```

Do you see a difference here?
Both pf_bh and pf_rebal contains useful information that we can extract.

```{r}
# Create eop_weight_bh
eop_weight_bh <- pf_bh$EOP.Weight

# Create eop_weight_rebal
eop_weight_rebal <- pf_rebal$EOP.Weight

# Plot end of period weights
par(mfrow = c(2, 1), mar=c(2, 4, 2, 2))
plot.zoo(eop_weight_bh$KO)
plot.zoo(eop_weight_rebal$KO)
```

We can see here that: 
1. The buy and hold strategy, the weights is changes due to the fluctuation of the prices. 
2. The rebalance strategy, it always try to rebalance it to the initial set weights.

Portfolio Return Analysis is both analyse the past performance and predict the future. How to choose the best portfolio?

We should always consider the risk vs reward. Portfolio management is heavily utilized math, statistics, programming skills and intuition.

How to quantify risk and reward? 
Reward -> portfolio mean return, it is the expectation return of a portfolio 
Risk -> portfolio volatility (variance and standard deviation), it is a measure how big the changes of prices againt the mean return

Return is not linear. 50% loss then followed by 50% profit, we only obtain our money 75% from the initial value. final value = initial value * 0.5 * 1.5 = 0.75 * initial value That’s why we won’t use simple arithmetic when calculate returns. 

Instead we will used what we called geometric return

Now let’s try to analyze the sp500 index.

```{r}
# Fetch price
sp500 <- getSymbols("^GSPC", from = "1985-01-01", to = "2017-12-30", src = "yahoo", auto.assign = FALSE)
head(sp500)
```

```{r}
# Convert the daily frequency of sp500 to monthly frequency: sp500_monthly
sp500_monthly <- to.monthly(sp500)
# Print the first six rows of sp500_monthly
head(sp500_monthly)
```

```{r}
# Create sp500_returns using Return.calculate using the closing prices
sp500_returns <- Return.calculate(Cl(sp500_monthly))
# Time series plot
plot.ts(sp500_returns)
```

```{r}
# Produce the year x month table
table.CalendarReturns(sp500_returns)
```

```{r}
# Remove first value because it is NA
sp500_returns <- sp500_returns[(-1),]
# Compute the mean monthly returns
mean(sp500_returns)
```


```{r}
# Compute the geometric mean of monthly returns
mean.geometric(sp500_returns)
```

```{r}
# Compute the standard deviation
sd(sp500_returns)
```

### The (annualized) Sharpe ratio 
Here we will define risk-free asset is something like money market fund or simply Bank Account. It is depends on you what you want to choose. Money market fund actually not 100% risk-free. Depends on the composition of the fund, if the risk is small enough, we can safely ignore it.

In this stage, you should already understand risk-return tradeoff and its graph representation. 

You will also want to learn about Modern Portfolio Theory, Excess Return Sharpe Ratio.

### Annualized mean and volatility

```{r}
table.AnnualizedReturns(sp500_returns)
```

### Time-variation in portfolio performance 
So far we only view everything as static. In reality, asset price is affected by things like: cycle, news, psychological effect, political, unexpected pandemic, etc. 

That’s why we need to weight more the most recent observation and weight less (or discard) the most distant, it is called: 
Rolling estimation. The tricky part here is to choose the window length? How should we choose the window length?

Rolling annualized mean and volatility

```{r}
# Plotting the 12-month rolling annualized mean
chart.RollingPerformance(R = sp500_returns, width = 12, FUN = "Return.annualized")
```

```{r}
# Plotting the 12-month rolling annualized standard deviation
chart.RollingPerformance(R = sp500_returns, width = 12, FUN = "StdDev.annualized")
```

```{r}
# Plotting the 12-month rolling annualized Sharpe ratio, assuming rf = 0
chart.RollingPerformance(R = sp500_returns, width = 12, FUN = "SharpeRatio.annualized")
```

To see all together:


```{r}
charts.RollingPerformance(R = sp500_returns, width = 12)
```

### Using window to analyze specific date range

```{r}
# Use daily return
sp500_ret_daily <- Return.calculate(Cl(sp500))

# Fill in window for 2008
sp500_2008 <- window(sp500_ret_daily, start = "2008-01-01", end = "2008-12-31")

# Create window for 2014
sp500_2014 <-window(sp500_ret_daily, start = "2014-01-01", end = "2014-12-31")

# Plotting settings
par(mfrow = c(1, 2) , mar=c(3, 2, 2, 2))
names(sp500_2008) <- "sp500_2008"
names(sp500_2014) <- "sp500_2014"

# Plot histogram of 2008
chart.Histogram(sp500_2008, methods = c("add.density", "add.normal"))

# Plot histogram of 2014
chart.Histogram(sp500_2014, methods = c("add.density", "add.normal"))
```

### How does it compare to normal distribution? 
Is it correct to assume asset return on normal distribution? Of course no. It is almost illogical to use normal distribution on asset return.

Non-normality of the return distribution Asset return tend to be skewed and fat-tailed. It is possible to see more negative returns on the left tail of the distribution.

Two metrics key to understanding the distribution of non-normal returns are skewness and kurtosis. 

```{r}
# Calculate both skewness and kurtosis in R are very easy
skewness(sp500_ret_daily)
```

```{r}
kurtosis(sp500_ret_daily)
```

### Semi-Deviation, Value-at-Risk and 5% Expected Shortfall In this stage, you should already understand the terms: 5% expected shortfall and Value-at-Risk 

Semi-deviation is a method of measuring the below-mean fluctuations in the returns on investment. 
```{r}
SemiDeviation(sp500_returns)
```

```{r}
VaR(sp500_returns, p = 0.05)
```

```{r}
ES(sp500_returns, p = 0.05)
```

### Drawdown 
We now analyzing the portfolio’s drawdowns, or peak-to-trough decline in cumulative returns. 

```{r}
table.Drawdowns(sp500_ret_daily)
```

```{r}
chart.Drawdown(sp500_ret_daily)
```


**The future are unknown and random ** 
We will start to shift, from thinking return as past performance, to return as predictions. In this stage, you should already understand the terms: 
Expected Value, Variance, and Random Variable. 

Correlation between returns We are using ko_ret and pep_ret again.

```{r}
# Create a scatter plot of returns
chart.Scatter(ko_ret, pep_ret, xlab = "ko returns", ylab = "pep returns", main = "ko-pep returns")

```

```{r}
# Find the correlation
cor(ko_ret, pep_ret)
```

```{r}
# Find and visualize the correlation using chart.Correlation
chart.Correlation(cbind(ko_ret, pep_ret))
```

```{r}
# Visualize the rolling estimates using chart.RollingCorrelation
chart.RollingCorrelation(ko_ret, pep_ret, width = 22) # 22 trading days, about 1 calendar month
```

```{r}
chart.RollingCorrelation(ko_ret, pep_ret, width = 252) # 252 trading days, about 1 calendar year

```

### Covariance Matrix

```{r}
ko <- getSymbols("ko", from = "2003-01-01", to = "2016-08-30", auto.assign = FALSE)
pep <- getSymbols("pep", from = "2003-01-01", to = "2016-08-30", auto.assign = FALSE)
aapl <- getSymbols("aapl", from = "2003-01-01", to = "2016-08-30", auto.assign = FALSE)
ge <- getSymbols("ge", from = "2003-01-01", to = "2016-08-30", auto.assign = FALSE)

```

```{r}
ret_ko <- Return.calculate(Ad(ko))
ret_pep <- Return.calculate(Ad(pep))
ret_aapl <- Return.calculate(Ad(aapl))
ret_ge  <- Return.calculate(Ad(ge))

```

```{r}
# Remove NA value in the first row
ret_ko <- ret_ko[(-1),]
ret_pep <- ret_pep[(-1),]
ret_aapl <- ret_aapl[(-1),]
ret_ge  <- ret_ge[(-1),]

```

```{r}
return_multi <- cbind(ret_ko, ret_pep, ret_aapl, ret_ge) # you could insert as many return as you like

# Create a vector of returns 
means <- apply(return_multi, 2, "mean")
  
# Create a vector of standard deviation
sds <- apply(return_multi, 2, "sd")

# Create a scatter plot
plot(sds, means)
text(sds, means, labels = colnames(return_multi), cex = 0.7)
abline(h = 0, lty = 3)

```


 Do you see any pattern/cluster here?

Now we will compute the covariance and correlation matrix on the daily returns of the four assets from the previous exercise.

```{r}
# Create a matrix with variances on the diagonal
diag_cov <- diag(sds^2)

# Create a covariance matrix of returns
cov_matrix <- cov(return_multi)

# Create a correlation matrix of returns
cor_matrix <- cor(return_multi)

# Verify covariances equal the product of standard deviations and correlation
all.equal(cov_matrix[1,2], as.numeric(cor_matrix[1,2] * sds[1] * sds[2]))

```

Portfolio risk budget
```{r}
# Create portfolio weights [arbitrary in this case, for practice purpose]
weights <- c(0.4, 0.4, 0.1, 0.1)
# Create volatility budget
vol_budget <- StdDev(return_multi, portfolio_method = "component", weights = weights)
# Make a table of weights and risk contribution
weights_percrisk <- cbind(weights, vol_budget$pct_contrib_StdDev)
colnames(weights_percrisk) <- c("weights", "perc vol contrib")
# Print the table
weights_percrisk

```


Modern Portfolio Theory (MPT)
NB: Post-modern portfolio theory (PMPT) attempts to improve on modern portfolio theory by minimizing downside risk instead of variance.

```{r}
tickers <- c("AA","AAPL","AXP","BA","BAC","CAT","CVX","DD","DIS","GE","HD","HPQ","INTC","IBM","JNJ","JPM","KO","MCD","MMM","MRK","MSFT","NKE","PFE","PG","TRV")
# ,"UTX","VZ","WMT","XOM","T"
first.date = "1991-01-01"
last.date = "2016-12-31"
getSymbols(tickers, from=first.date, to=last.date, periodicity = 'monthly')

```

```{r}
ClosePrices <- lapply(tickers, function(x) Ad(get(x)))
```

```{r}
# The process to combine into one dataframe a bit messy here, there should be a better way
dim2 <- dim(ClosePrices[[1]])[1]
df_fin = data.frame(matrix(nrow = dim2))
for (i in 1:length(ClosePrices)){
  df_fin <- cbind2(df_fin, ClosePrices[i])
}
df_fin <- df_fin[,(-1)]
head(df_fin)
```


```{r}
stocks_ret <- CalculateReturns(as.xts(df_fin))
# Drop first row because it is NA
stocks_ret <- stocks_ret[(-1),]
head(stocks_ret)
```

Finding the mean-variance efficient portfolio 
We will use portfolio.optim function from tseries https://www.rdocumentation.org/packages/tseries/versions/0.10-47/topics/portfolio.optim The only argument needed is the monthly return data on the portfolio components for which the weights need to be determined.

```{r}
# Load tseries
library(tseries)
# Create an optimized portfolio of returns
opt <- portfolio.optim(stocks_ret)

# Create pf_weights
pf_weights <- opt$pw

# Assign asset names
names(pf_weights) <- colnames(stocks_ret)

# Select optimum weights opt_weights
opt_weights <- pf_weights[pf_weights >= 0.01]

# Bar plot of opt_weights
barplot(opt_weights)
```

```{r}
# Print expected portfolio return
opt$pm

```


```{r}
# Print expected portfolio volatility
opt$ps

```

The argument that can be supplied into the portfolio.optim() are: pm: the desired mean portfolio return. 
riskless: a logical indicating whether there is a riskless lending and borrowing rate. 
shorts: a logical indicating whether shortsales on the risky securities are allowed. 
rf: the risfree interest rate. etc…..

This exercise will show the effect of increasing your target return on the volatility of your mean-variance efficient portfolio.
```{r}
# Create portfolio with target return of average returns 
pf_mean <- portfolio.optim(stocks_ret, pm = mean(stocks_ret))

# Create portfolio with target return 10% greater than average returns
pf_10plus <- portfolio.optim(stocks_ret, pm = 1.1 * mean(stocks_ret))

# Print the standard deviations of both portfolios
pf_mean$ps
```

```{r}
pf_10plus$ps

```

```{r}
# Calculate the proportion increase in standard deviation
(pf_10plus$ps - pf_mean$ps) / (pf_mean$ps)

```

In-sample vs. out-of-sample evaluation Split the dataset into two parts: 
1. Estimation sample: used to find the optimal weights 
2. Out of sample: evaluation to give a realistic view on portfolio performance

### Split-sample evaluation

```{r}
# Create returns_estim 
returns_estim <- window(stocks_ret, end = "2003-12-31")

# Create returns_eval
returns_eval <- window(stocks_ret, start = "2004-01-01")

# Create vector of max weights
max_weights <- rep(0.1, ncol(stocks_ret))

# Create portfolio with estimation sample 
pf_estim <- portfolio.optim(returns_estim, reshigh = max_weights)

# Create portfolio with evaluation sample
pf_eval <- portfolio.optim(returns_eval, reshigh = max_weights)

# Create a scatter plot with evaluation portfolio weights on the vertical axis
plot(pf_estim$pw, pf_eval$pw)
abline(a = 0, b = 1, lty = 3)
```

By creating two samples using the window() function you can backtest, or evaluate the performance of a portfolio on historical data!

Out of sample performance evaluation This example will illustrate how your returns can change based on the weighting created by an optimized portfolio. You will use the estimation portfolio (pf_estim) to evaluate the performance of your portfolio on the estimation sample of returns (returns_eval).

How severe is the optimality loss? 
Let’s compare, for the portfolio weights in pf_estim, the performance you expected using the evaluation sample (returns_estim) with the actual return on the out-of-sample period (returns_eval).

```{r}
# Create returns_pf_estim
returns_pf_estim <- Return.portfolio(returns_estim, pf_estim$pw, rebalance_on = "months")

# Create returns_pf_eval
returns_pf_eval <- Return.portfolio(returns_eval, pf_estim$pw, rebalance_on = "months")

```

```{r}
# Print a table for your estimation portfolio
table.AnnualizedReturns(returns_pf_estim)
```

```{r}
# Print a table for your evaluation portfolio
table.AnnualizedReturns(returns_pf_eval)

```

 
A little bit more sophisticated Read the paper here: https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Ardia~et~al.pdf You may also want to learn more about the package used in this section: https://cran.r-project.org/web/packages/PortfolioAnalytics/vignettes/portfolio_vignette.pdf
