---
title: "QRM P01"
author: "Alexander McNeil"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Alex McNeil is author of Quantitative Risk Management: Concepts, Techniques and Tools.
The associated tutorial is on the website [qrmtutorial](https://qrmtutorial.org/)

We use two packages here:
1. qrmdata: financial dataset
2. qrmtools: Necessary functions

### Why QRM?
In QRM, the goal is to quantify the risk of a portfolio of risky assets. Measuring the risk is the first step towards managing the risk. we can consider the typical portfolios of risky assets held by banks and insurance companies, and sometimes also their liabilities. But the ideas apply equally to the portfolio of a private investor. 

Managing the risk can entail many things. You might try to reduce the risk by selling assets, by acquiring new assets to increase the diversification of the portfolio, or by using so-called hedging strategies. 

For the banks and insurers, an important part of managing risk is making sure they have sufficient capital to withstand large losses on their portfolios and remain solvent. 

For this purpose, they compute measurements of the amounts that they could lose in periods of extreme market stress. 

A well-known example of one of these risk measures is Value-at-Risk. You will work up to calculating Value-at-Risk for portfolios. 

#### Risk factors

Think of a portfolio of risky assets. It might contain some stock index trackers, some individual stocks, some government bonds or treasuries, some corporate bonds. It might contain assets denominated in domestic currency and some denominated in foreign currencies. It might contain commodities, like gold. It might contain some derivative securities designed to hedge risk or speculate on price movements, such as equity options. 

The value of this portfolio at any point in time depends on the fluctuating values of many underlying quantities, which we call risk factors. Examples of risk factors are equity indexes, individual equity prices, foreign exchange rates, interest rates for different borrowing periods (in the case of bonds), and commodity prices. 

Let's load the QRMdata package and then the dataset SP500. This contains daily values of the well-known equity index based on 500 of the most important US stocks.
### Analyzing risk factors with R
```{r}
library(qrmdata)
data(SP500)

head(SP500, n = 3)
tail(SP500, n = 3)

```

### Plotting risk factors
```{r}
plot(SP500)
```

## Exploring risk-factor time series: equity indexes

In this exercise, you will look at an equity index and plot it for a particular range of dates. The data used in this exercise and in the rest of the course are contained in the package qrmdata. You also need the package xts to manipulate time series.

When the qrmdata library is attached, as it will be throughout the course, you can load a dataset with the data() command. For example, the command data("FTSE") loads the UK FTSE (Financial Times Stock Exchange) index, which you can then refer to as object FTSE.

If you want to extract the data from a certain date range, for example from April 1 to June 30, 2000, you can create a new object using the command ftse00 <- FTSE["2000-04-01/2000-06-30"].


    Load the Dow Jones index "DJ" from qrmdata.
    Show the first and last few lines of the DJ index with head() and tail().
    Plot the DJ index using plot().
    Extract the DJ index for the crisis period 2008-2009 and assign to object dj0809.
    Plot the dj0809 using the same plotting function as above.

```{r}
# Load DJ index
data()

# Show head() and tail() of DJ index

# Plot DJ index

# Extract 2008-2009 and assign to dj0809

# Plot dj0809
```

## Exploring risk-factor time series: individual equities

For some risk management applications, it is sufficient to model equity risk by looking at indexes. If you want a more detailed model of the risk in a portfolio of equities, you can drill down to the level of individual share prices.

In the previous chapter, you used DJ["2008/2009"] to extract the Dow Jones data from certain rows of an xts object by specifying a date range index. To also extract data from particular columns, you can add a column identifier, like a string name or numeric index, in the brackets following a comma. To select multiple columns, include these column identifiers in a vector. This [rows, columns] format is consistent with indexing most other two dimensional objects in R.

data[index, colname]
data[index, c(col1index, col2index)]

The qrmdata package also includes data for certain constituents, or the stocks or companies part of a larger index. The Dow Jones constituents data are contained in "DJ_const". In this exercise, you will view the names of all its stocks, select the Apple and Goldman Sachs share prices, and plot them using the command plot.zoo() to display multiple time series.
Instructions
100 XP

    Load the DJ constituents data "DJ_const" from qrmdata.
    Use names() to view the names in DJ_const and head() to display the first few rows.
    Extract only the Apple ("AAPL") and Goldman Sachs ("GS") share prices for 2008-2009 and assign them to object stocks.
    Plot stocks using plot.zoo().

```{r}
library(zoo)
# Load DJ constituents data
data(DJ_const)

# Apply names() and head() to DJ_const

# Extract AAPL and GS in 2008-09 and assign to stocks
stocks <- 

# Plot stocks with plot.zoo()
```

## Exploring risk-factor data: exchange rates

For a portfolio with risk exposure in different countries, it is necessary to consider the risk coming from foreign exchange (FX) rates. The qrmdata package includes FX rate data for many currencies, ranging from Swiss Francs to Japanese Yen, with respect to the USD (United States dollar) and GBP (Great Britain pound).

In this exercise, you will look at the datasets "EUR_USD" and "GBP_USD", which contain the Euro and British pound exchange rates against the US dollar. Then, you will merge these time series and plot them together for the period 2010-2015.
Instructions
100 XP

    Load the foreign exchange data "GBP_USD" and "EUR_USD" from qrmdata.
    Use plot() to plot each exchange rate separately.
    Use plot() and the inverse of GBP_USD to plot a US dollar to British pound exchange rate.
    Use merge() to merge the GBP_USD and EUR_USD data, in that order, as object fx.
    Extract the exchange rates for 2010-15 from fx and assign to fx0015.
    Plot fx0015 using plot.zoo().

```{r}
# Load exchange rate data
data("GBP_USD")
data("EUR_USD")

# Plot the two exchange rates

# Plot a USD_GBP exchange rate

# Merge the two exchange rates GBP_USD and EUR_USD
fx <- 

# Extract 2010-15 data from fx and assign to fx0015
fx0015 <- 

# Plot the exchange rates in fx0015
```

# 1. Risk-factor returns

In QRM, the aim is to model the fluctuations in key risk factors which affect the value of a portfolio.

Risk-factor returns

These fluctuations are called the risk-factor changes or risk-factor returns or simply returns. However, there are a few different ways of defining returns. Let (Z_t) be a time series containing the values of a risk factor at a time at a set of regularly-spaced times, which could represent days, weeks, months, etc. For illustration, let's suppose it is a daily series. Here are three different ways of defining risk-factor returns (X_t): In the first definition, the returns are the differences of the risk-factor values, known as simple returns. This is the simplest definition but not the most common. It tends to be used when the risk factors have very small values close to zero (like certain interest-rate series). The second definition is the easiest to interpret. The relative returns are the differences divided by the initial values. If you multiply the relative returns by one hundred, you get percentage changes. So if a stock has a relative return of point zero two, it gains 2% in value; if it has a relative return of minus point zero three, it falls 3% in value. In the third definition, the returns are the differences of the log-values of the risk factors; these are log-returns. This definition is, in fact, the most widely used. Here are a few of the reasons why log-returns are popular.

Properties of log-returns

If you build a model for the log-returns of a risk factor, you know that the risk-factor can never become negative. This is generally a desirable feature for risk factors that are prices and rates. However, some risk factors can become negative under unusual market conditions, an example being short-term interest rates. Log returns are, in fact, very close to relative returns for typical values. It can be shown with some algebra that log of Z_t minus log of Z_(t-1) can be approximated by Z_t minus Z_(t-1) divided by Z_(t-1) Finally, there are a couple of theoretical reasons for using log-returns: It is particularly easy to aggregate them to calculate longer-interval returns, such as weekly or monthly returns. Also, there is a very famous model for prices known as the Black-Scholes-Merton model, or geometric Brownian motion (GBM). The model is widely used for pricing derivatives. If a price series follows GBM, then the log-returns will be normally distributed. That's all you really need to know about this model at this stage.


Log-returns in R

It's worth remarking, however, that you are going to find that log-returns generally aren't normally distributed in practice, at least for short time horizons. To calculate the log-returns of the S&P 500 index, you apply the functions log and diff in succession. Diff calculates differences. Here is some code to illustrate. Note that you get an NA in the first position when you apply these operations. It is best to remove the first value prior to working with the return series.

```{r}
sp500x <- diff(log(SP500))[-1]
head(sp500x)
```

```{r}
plot(sp500x)
```

Log-returns in R (2)

Now let's plot the log-returns. You see, a fairly noisy series mostly oscillating between minus point zero five and plus point zero five, which can be approximately interpreted as 5% movements up and down. However, there are some very extreme values. The largest negative return corresponds to the famous stock market crash on Black Monday in 1987.


## Exploring return series

To analyze risk, the key task is to model the fluctuations in prices and rates over different time periods; these fluctuations are known as returns. To calculate the log-returns of the FTSE stock index and assign to ftse_x, apply the log() and diff() functions in succession:

```{r}
data("FTSE")
ftse_x <- diff(log(FTSE))
```

As you saw in the video, differencing in this way will always give a NA in the first position of the time series, which can then be removed with diff(log(FTSE))[-1]. However, you will not need to do this in the course unless it is specified in the instructions.

In this exercise, you will calculate and plot log-return series for the equity and FX risk factors that you have previously encountered. The datasets dj0809, djstocks, and GBP_USD have been pre-loaded into your workspace.
Instructions
100 XP

    Compute the log-returns of the DJ index in dj0809 and assign to object dj0809_x.
    Plot the return series dj0809_x.
    Compute the log-returns of all share prices in djstocks and assign to djstocks_x.
    Plot the share returns djstocks_x. Note that djstocks_x contains multiple time series.
    Compute the log-returns of the GBP_USD exchange rate series and assign to erate_x.
    Plot the return series erate_x.

```{r}
library(qrmdata)
library(qrmtools)
data(djstocks)
# Compute the log-returns of dj0809 and assign to dj0809_x
dj0809_x <- 

# Plot the log-returns

```
```{r}
# Compute the log-returns of djstocks and assign to djstocks_x
djstocks <- DJ_const["2008/2009", c("AAPL", "GS")]
djstocks_x <- 

# Plot the two share returns

```

```{r}
# Compute the log-returns of GBP_USD and assign to erate_x
erate_x <- 

# Plot the log-returns
```
The return series often just look like noise with some periods of larger fluctuations. You’ll discover later that they typically have a lot of interesting structure.

## Different ways of plotting risk-factor and return series

You already know that you can use plot.zoo() to plot multiple time series. For a four-dimensional time series data, the call plot.zoo(data) creates four separate plots by default, unless you include the parameter plot.type = “single” to plot all four series in one plot. You can also add even more parameters such as col to specify different colors and type = “h” to get vertical bars instead of joining points, which can sometimes be a better way of displaying returns.

plot.zoo(x, plot.type, col = 1, type = "l", ...)

In this exercise, you will explore the plot.zoo() function to plot equity risk-factor data and the corresponding returns in different ways.
Instructions


    Plot djstocks in four separate plots.
    Plot djstocks in one plot in colors 1 to 4. The code to create an appropriate legend for the plot is provided.
    Compute the log-returns of djstocks and assign them to djstocks_x.
    Plot djstocks_x in four separate plots.
    Plot djstocks_x in four separate plots with vertical bars.


```{r}
djstocks <- DJ_const["2008/2009", c("AAPL", "AXP", "BA", "CAT")]
# Plot djstocks in four separate plots


# Plot djstocks in one plot and add legend
plot.zoo(djstocks, plot.type = "single", col = c(1:4))
legend(julian(x = as.Date("2009-01-01")), y = 70, legend = names(DJ_const)[1:4], fill = 1:4)


# Compute log-returns and assign to djstocks_x
djstocks_x <- 

# Plot djstocks_x in four separate plots


# Plot djstocks_x with vertical bars

```


Note how in late 2008 there were large returns for all series. That was the height of the financial crisis.


## Aggregating log-return series
In statistics, aggregate data are data combined from several measurements. You just learned that you can compute compute weekly, monthly and quarterly log-returns by summing daily log-returns with the corresponding apply.weekly(), apply.monthly() and apply.quarterly() functions.

For example, you can use the following code to form the quarterly returns for a univariate time series data and multivariate time series mv_data:

# apply.quarterly(x, FUN, ...)
data_q = apply.quarterly(data, sum)
mv_data_q = apply.quarterly(mv_data, colSums)

In this exercise, you will practice aggregating time series data using these functions and plotting the results. The data DJ and DJ_const are available in your workspace, as are the objects djx, which contains daily log-returns of the Dow Jones index from 2000-2015, and djreturns, which contains the daily log-returns for the first four DJ_const stocks from 2000-2015. Use plot for univariate time series and plot.zoo for multivariate time series.

Instructions

    Plot the object djx.
    In one line, plot the weekly log-returns of djx with vertical bars.
    Plot the monthly log-returns of djx with vertical bars.
    Plot the object djreturns using plot.zoo.
    Plot the monthly log-returns for djreturns with vertical bars using plot.zoo.

```{r}
dj <- DJ["2000/2015"]
djx <- diff(log(dj))
djstocks <- DJ_const["2000/2015",c("AAPL", "AXP", "BA", "CAT")]
djreturns <- diff(log(djstocks))

# Plot djx


# Plot weekly log-returns of djx


# Plot monthly log-returns of djx


# Plot djreturns


# Plot monthly log-returns of djreturns
```

These aggregation functions are extremely useful as for analyzing risk over longer time horizons.

## A test on aggregation of log-returns
Data scientists often use the aggregations that you have learned so far in combination with summary statistics to extract even more insights from data. Functions that calculate summary statistics include mean(), median(), and var(). The object sp contains daily log-returns for the S&P 500 index for the period 1960-2015; it is loaded in your workspace. To three decimal places, what is the average quarterly log-return for the S&P 500 from 1990-2010?

```{r}
data("SP500")
sp <- SP500["1990/2010"]
sp <- diff(log(sp))[-1]
mean(apply.quarterly(sp, sum))
```

## Exploring other kinds of risk factors

The plotting function pairs() creates a pairwise scatterplot of the components of a multivariate time series with two or more dimensions. It is used on a zoo object rather than an xts object.

A roughly circular shape of a scatterplot indicates a low correlation between the log-returns of two different commodities. Generally speaking, low correlation is good in a portfolio as it implies that the assets are diversified. High correlation, on the other hand, represents a risk that must be properly modelled.

In this exercise, you will look at gold and oil prices over a 25 year period, calculate their daily and monthly log-returns, and plot them. The data gold and oil, containing the daily prices from 1990-2015 of gold and Brent crude oil.

Instructions

    Use plot() to plot the gold and oil time series separately.
    Calculate the daily log-returns of each commodity and assign to goldx and oilx, respectively.
    Calculate the monthly log-returns of each commodity and assign to goldx_m and oilx_m, respectively.
    Use merge() to merge goldx_m and oilx_m, in that order, into coms.
    Plot coms, a multivariate series, with vertical bars.
    Convert coms to a zoo object with as.zoo() and then apply pairs() to create a pairwise scatterplot.


```{r}
data("OIL_Brent")
oil <- OIL_Brent["1990/2015"]
data("GOLD")
gold <- GOLD["1990/2015"]

# Plot gold and oil prices



# Calculate daily log-returns
goldx <- 
oilx <- 

# Calculate monthly log-returns
goldx_m <- apply.monthly(goldx, sum)
oilx_m <- apply.monthly(oilx, sum)

# Merge goldx_m and oilx_m into coms
coms <- merge(goldx_m, oilx_m)

# Plot coms with vertical bars


# Make a pairwise scatterplot of coms
```


As you can see, gold and oil are well diversified commodities.

## Interest-rate data

The object zcb contains daily values of Canadian zero-coupon-bond yields, expressed as percentages, for the period 2006-2015. Yields are the key risk-factor when it comes to analysing the interest-rate risk in a portfolio of bonds or other fixed-income products.

It is not so clear what is the best way of calculating risk-factor changes for yields. It is possible to compute log-returns, provided yields are not negative, and it is also possible to calculate simple returns. To compute the simple returns of a series, use only diff() instead of diff() and log().

In this exercise, you will plot time series of yields for fixed times to maturity, and plot risk-factor changes for these yields. You will also plot the whole yield curve on particular dates. The zcb data has been loaded into your workspace. A vector yield_cols containing the names of the columns corresponding to maturities of 1, 5 and 10 years has been created. A numerical vector maturity containing all the maturities in years has also been created.

Instructions
100 XP

    Compute the log-returns of zcb as zcb_x and the simple log-returns as zcb_x2.
    Plot zcb_x for 1, 5 and 10-year maturities in one plot.
    Plot zcb_x2 for 1, 5 and 10-year maturities in one plot.
    Index zcb in plot() to plot the yield curve for the first day in zcb.
    Index zcb in lines() to add a line to the yield curve for the last day in zcb.

```{r}
data("ZCB_CAD")
zcb <- ZCB_CAD["2006/2015"]

maturity <- c(0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 2.25, 2.5, 2.75, 3, 
3.25, 3.5, 3.75, 4, 4.25, 4.5, 4.75, 5, 5.25, 5.5, 5.75, 6, 6.25, 
6.5, 6.75, 7, 7.25, 7.5, 7.75, 8, 8.25, 8.5, 8.75, 9, 9.25, 9.5, 
9.75, 10, 10.25, 10.5, 10.75, 11, 11.25, 11.5, 11.75, 12, 12.25, 
12.5, 12.75, 13, 13.25, 13.5, 13.75, 14, 14.25, 14.5, 14.75, 
15, 15.25, 15.5, 15.75, 16, 16.25, 16.5, 16.75, 17, 17.25, 17.5, 
17.75, 18, 18.25, 18.5, 18.75, 19, 19.25, 19.5, 19.75, 20, 20.25, 
20.5, 20.75, 21, 21.25, 21.5, 21.75, 22, 22.25, 22.5, 22.75, 
23, 23.25, 23.5, 23.75, 24, 24.25, 24.5, 24.75, 25, 25.25, 25.5, 
25.75, 26, 26.25, 26.5, 26.75, 27, 27.25, 27.5, 27.75, 28, 28.25, 
28.5, 28.75, 29, 29.25, 29.5, 29.75, 30)

yield_cols <- c("1.00y", "5.00y", "10.00y")
# Compute log-returns as zcb_x and simple returns as zcb_x2
zcb_x <- 
zcb_x2 <- 

# Plot zcb_x for 1, 5 and 10-year maturities


# Plot zcb_x2 for 1, 5 and 10-year maturities


# Plot the yield curve for the first day of zcb
plot(maturity, zcb[1,], ylim = range(zcb), type = "l", ylab = "yield (%)", col = "red")

# Add a line for the last day of zcb
lines(maturity, zcb[nrow(zcb), ])
```

Yields can seem a little tricky to work with at first but they are fundamental for analyzing bond portfolios and many other financial products that depend on interest rates.